\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-31 background.tex
Updated
Updated the paper

SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
running at their individual pace, interacting with the environment via
signals and with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adhere to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, execution of every
statement in the logical zero time requires $\delta$ physical time. Time
for a logical tick can thus be summarized as $\Delta = \mathcal{F}
(\delta)$, where $\mathcal{F}$ is the schedule. We call $\Delta$ the
reaction time. Depending upon the amount of computation required and
schedule, $\Delta$ can vary. In order to satisfy the implicit
restriction imposed by the synchrony hypothesis -- no input event
\textit{from the environment} can be missed -- one needs to calculate
the \textit{Worst Case Reaction Time} (WCRT) and the resultant WCRT
needs to be smaller than the time between any two consecutive events on
any input, else the synchrony hypothesis is violated. Techniques for
calculating the WCRT of synchronous programs~\cite{boldt07} exist and
this analysis is not the focus of this paper. The opposite of the WCRT,
the shortest time for a reaction, is called the \textit{Best Case
  Reaction Time} (BCRT). Formally, let $\{\Delta_1, \Delta_2,\ldots,
\Delta_i\}$, be the set of all possible reaction times for some
CD. Then, $WCRT= max (\Delta_i)$ and $BCRT = min (\Delta_i)$.

>>>>

SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
running at their individual pace, interacting with the environment via
signals and with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs are composed
of one or more units called reactions. These reactions themselves can be
composed of children reaction forming a hierarchy. Each CD adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous subset of SystemJ. Unfortunately, execution of every
statement in the logical zero time requires $\delta$ physical time. Time
for a logical tick can thus be summarized as $\Delta = \mathcal{F}
(\delta)$, where $\mathcal{F}$ is the schedule. We call $\Delta$ the
reaction time. Depending upon the amount of computation required and
schedule, $\Delta$ can vary. In order to satisfy the implicit
restriction imposed by the synchrony hypothesis -- no input event
\textit{from the environment} can be missed -- one needs to calculate
the \textit{Worst Case Reaction Time} (WCRT) and the resultant WCRT
needs to be smaller than the time between any two consecutive events on
any input, else the synchrony hypothesis is violated. Techniques for
calculating the WCRT of synchronous programs~\cite{boldt07} exist and
WCRT analysis is not the focus of this paper. The opposite of the WCRT,
the shortest time for a reaction, is called the \textit{Best Case
  Reaction Time} (BCRT). Formally, let $\{\Delta_1, \Delta_2,\ldots,
\Delta_N\}$, be the set of all possible reaction times for some
CD. Then, $\exists i \in N, \mathrm{where\ } WCRT= max (\Delta_i)$,
similarly, $\exists j \in N$, where $BCRT = min (\Delta_j)$.

### 2014-01-31 correct.tex
Updated the paper

Next, we prove the completeness property.
\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for
  \textbf{all} given reaction times $\Delta$ in the range of $BCRT \leq
  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
\end{theorem}
\begin{proof}
  From Theorem~\ref{th:1} case 3.
\end{proof}

@@

The proofs for correctness, completeness, and soundness for
Algorithm~\ref{alg:2} are similar to those of
Algorithm~\ref{alg:1}. Furthermore, the proofs for \texttt{wait\_atleast
  (M)} and \texttt{wait\_exact (M)} can be proven using the same
techniques as shown in Theorem~\ref{th:1}.

>>>>

Next, we prove the completeness property.
\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for
  \textbf{all} given reaction times $\Delta$ in the range of $BCRT \leq
  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
\end{theorem}
\begin{proof}
  Follows from Theorem~\ref{th:1}.
\end{proof}

@@

The proofs for correctness, completeness, and soundness for
Algorithm~\ref{alg:2} are similar to those of
Algorithm~\ref{alg:1}. Furthermore, even correctness, completeness, and
soundness for \texttt{wait\_atleast (M)} and \texttt{wait\_exact (M)}
can be proven using the same techniques as shown in Theorem~\ref{th:1}.

### 2014-01-31 delays.tex
Updated the paper

Given a SystemJ program: \texttt{wait\_inbetween (M..N); p}, where $M
\in \mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is
executed after real-time delay of $\tau$ units, such that, $M \leq \tau
\leq N$.

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} $M$ times
  units.
\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

@@

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps the
physical notion of time back to the elegant logical notion of time. The
rewrite \textit{waits} a certain number of logical ticks, before
proceeding to the next statement. The number of logical ticks \texttt{d}
to \textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
statically at compile time. The value of \texttt{d} is intricately tied
with the WCRT and BCRT of the program and hence the execution platform.

@@

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time wait
period. Figure~\ref{fig:hrtcssoln} gives the solution space for the
aforementioned calculation. The patterned space is the only overlapping
region and hence gives the solution space for the above example. It is
possible that $D$ is an empty set, thereby, no solution can be found. We
provide solutions for such cases in the next section.

@@

The \texttt{wait\_exact(M)} variant is more interesting. Like before,
we find sets $S_1$ and $S_2$, and find the intersection of the two sets
to get the value of \texttt{d}. It is possible % (and often likely, as
% suggested by our experiments in
% Section~\ref{sec:experimental-results})
that the resultant set $D$ is empty (also possible in the case of
\texttt{wait\_inbetween(M..N)}, but never possible in case of
\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
sets and will always contain at minimum a single element. For such cases
we provide two solutions listed below.

@@

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

>>>>

Given a  SystemJ program:  \texttt{wait\_inbetween (M..N); p},  where $M
\in  \mathbb{Q}^{>0}$  and $N  \in  \mathbb{Q}^{>0}$,  statement $p$  is
executed after  real-time postponement  of $\tau$  units, such  that, $M
\leq \tau \leq N$.

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time
  postponement of $\tau$ units, such that, $M \leq \tau < \infty$. It is
  important to note that the lower bound of the wait construct; $M$ is
  \textit{not} an exact postponement, but rather the control is allowed
  to proceed to the next statement anytime after waiting for \textit{at
    least} $M$ times units.
\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time
  postponement of $\tau$ units, such that, $M \leq \tau \leq M$. In this
  variant, the waiting time is exact.
\end{enumerate*}

@@

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps
real-time back to the elegant logical notion of time. The rewrite
\textit{waits} a certain number of logical ticks, before proceeding to
the next statement. The number of logical ticks \texttt{d} to
\textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
statically at compile time. The value of \texttt{d} is intricately tied
with the WCRT and BCRT of the program and hence the execution platform.

@@

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time wait
period. Figure~\ref{fig:hrtcssoln} gives the solution space for the
aforementioned calculation. The patterned space is the only overlapping
region and hence gives the solution space for the above example. It is
possible that $D$ is an empty set, that means no overlapping region
could be found. We provide solutions for such cases in the next
subsection.

@@

The \texttt{wait\_exact(M)} variant is more interesting. Like before,
we find sets $S_1$ and $S_2$, and find the intersection of the two sets
to get the value of \texttt{d}. It is possible % (and often likely, as
% suggested by our experiments in
% Section~\ref{sec:experimental-results})
that the resultant set $D$ is empty (also possible in the case of
\texttt{wait\_inbetween(M..N)}, but never possible in case of
\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
sets and will always contain at minimum a single element. For such
cases, we provide two solutions listed below.

@@

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, or is unable to provide a
postponement in the region of $[M, N]$ time units in the case of
\texttt{wait\_inbetween (M..N)} statement, thereby violating their
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD periodically at the WCRT, we can choose the value of \textit{d}
from set $S1$. This technique allows meeting the synchronous hypothesis,
while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous option,
this technique is very inflexible and results in reduced response time
and increased power consumption.

### 2014-01-31 discussion.tex
Updated the paper

We dedicate this section to discuss the details of the wait construct
semantics, their usage and the advantages and limitations of our
proposed technique to introducing these real-time constructs in a
reactive setting.

@@

\begin{figure}[h!]
	\vspace{-10pt}
        \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
while(true){wait_exact (1 ms); 
	emit S; //do something
}
 \end{lstlisting}
  \caption{Programming periodic tasks with wait\_exact statement}
  \label{fig:periodic}
	\vspace{-10pt}
\end{figure}

The program in Figure~\ref{fig:preemp}, reproduced from
Section~\ref{sec:progr-using-exact}, is supposed to emit signal
\texttt{S} every \texttt{1 ms}. Such periodic reactions (or CDs) require
special consideration, like any other real-time periodic task. One
should be aware of the interaction of wait and other \texttt{pause}
constructs in the body of the reaction (or CD). For example, consider
the (perceived) periodic program and its rewrite below.

@@

The system designer needs to be aware that just introducing a wait
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks, consequently increasing the
real-time wait period. Static analysis techniques from real-time
community can be used to automatically determine the wait specification,
in such cases, but describing these techniques is out of the scope of
this work.

@@

The proposed solution relies on time analyzable platforms, i.e.,
platforms where WCRT and BCRT can be calculated using static analysis
tools. Many might consider this to be too much of a restriction. In this
section we debunk this \textit{perceived} disadvantage.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The program in Figure~\ref{fig:preemp} sends a
signal to an external timer and waits for 10 ms to pass by. Consider
what happens when signals \texttt{S} and \texttt{TIMER} occur in the
same logical tick, the \texttt{await} statement is never executed (due
to suspend) and hence, we enter a deadlock. Such problems are completely
avoided in our technique, because we convert the real-time wait
constructs into logical waits (\texttt{pause} constructs), which
interact well with preemption.  Furthermore, synchronous/GALS languages
have always been targeted at real-time analyzable platforms and
constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
our technique within such a setting with ease.

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
		\begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
{
 //reaction R1
 emit START_TIMER(1 ms); 
 await (TIMER);emit O1;
}
|| //synchronous parallel
{
 //reaction R2
 // do something
}
		\end{lstlisting}
  \caption{Resolution of external timers}
  \label{fig:resolution}
	\vspace{-10pt}
\end{figure}

There is yet another problem with timer based systems. Let us consider
the program in Figure~\ref{fig:resolution}. On a cursory look this
example should work fine. An external timer is started in reaction R1,
which counts down from 1 ms. Once this time has elapsed, a signal
\texttt{TIMER} is generated from this external timer, which should emit
O1 in turn. There is no \texttt{suspend} construct encapsulating the
timer and hence all seems fine. Let us now consider what happens to the
whole CD, including reaction R2; while reaction R1 is waiting for the
\texttt{TIMER} signal, reaction R2 is making progress. Input signals can
only be captured from the environment at the logical tick boundaries
(cf. Section~\ref{sec:mapping-logical-time}). In
Figure~\ref{fig:resolution}, reaction R2 determines the length of the
tick, because it is performing the heavier computation. Consider what
happens if the external timer generates the \texttt{TIMER} signal, but
the system is still performing computation and the logical tick has not
yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
environment. Thus, even in the absence of the \texttt{suspend}
construct, we are not guaranteed that the program will capture the
external timer signal. In fact, the WCRT needs to be smaller than 1 ms
for the above program to perform as expected. But, this again implies
need for a time analyzable platform even when using external timers.

In general we observe that the WCRT determines the lowest resolution of
any real-time construct whether it be external timer independent, like
the proposed technique, or external timer dependent, like above. This
discovery is one of the key insights for proposing the solutions
described in this paper.

>>>>

We dedicate this section to discuss the details of the real-time wait
construct semantics, its advantages and limitations in a reactive
setting.

@@

% \begin{figure}[h!]
% 	\vspace{-10pt}
%         \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
% while(true){wait_exact (1 ms); 
% 	emit S; //do something
% }
%  \end{lstlisting}
%   \caption{Programming periodic tasks with wait\_exact statement}
%   \label{fig:periodic}
% 	\vspace{-10pt}
% \end{figure}

% The program in Figure~\ref{fig:preemp}, reproduced from
% Section~\ref{sec:progr-using-exact}, is supposed to emit signal
% \texttt{S} every \texttt{1 ms}.

One should be aware of the interaction of wait and other \texttt{pause}
constructs in the body of the reaction (or CD). For example, consider
the (perceived) periodic code snippet in Figure~\ref{pp:a} and its
rewrite in Figure~\ref{pp:b}.

@@

The system designer needs to be aware that just introducing a wait
construct does not make the reaction (or CD) periodic. The code snippet
in Figure~\ref{pp:a} does \textit{not} emit signal \texttt{S} every
\texttt{1 ms}, because of the extra \texttt{pause}
statement. Introducing additional pauses introduces more logical ticks,
consequently increasing the real-time wait period. Static analysis
techniques from real-time community can be used to automatically
determine the specification for the \texttt{wait\_exact} statement in
Figure~\ref{pp:a}, in such cases, but describing these techniques is out
of the scope of this work.

@@

The solution for introducing real-time as first class language
constructs introduced in this paper rely on platforms that allow
computing the WCRT and the BCRT of the program. Many might consider this
to be too much of a restriction. In this section we debunk this
\textit{perceived} disadvantage.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The program in Figure~\ref{fig:preemp} sends a
signal to an external timer and waits for 10 ms to pass by. Consider
what happens when signals \texttt{S} and \texttt{TIMER} occur in the
same logical tick, the \texttt{await} statement is never executed (due
to suspend) and hence, we enter a deadlock. Such problems are completely
avoided in our technique, because we convert the real-time wait
constructs into logical waits (\texttt{pause} constructs), which
interact well with preemption.  Furthermore, synchronous languages have
always been targeted at real-time analyzable platforms and constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
our technique within such a setting with ease.

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
        \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
{
 //reaction R1
 emit START_TIMER(1 ms); 
 await (TIMER);emit O1;
}
|| //synchronous parallel
{
 //reaction R2
 // do something with heavy computation
}
		\end{lstlisting}
  \caption{Resolution of external timers}
  \label{fig:resolution}
	\vspace{-10pt}
\end{figure}

There is yet another problem with timer based systems. Let us consider
the code snippet in Figure~\ref{fig:resolution}. On a cursory look this
example should work fine. An external timer is started in reaction R1,
which counts down from 1 ms. Once this time has elapsed, a signal
\texttt{TIMER} is generated from this external timer, which should emit
O1 in turn. There is no \texttt{suspend} construct encapsulating the
timer and hence all seems fine. Let us now consider what happens to the
whole CD, including reaction R2; while reaction R1 is waiting for the
\texttt{TIMER} signal, reaction R2 is making progress. Input signals can
only be captured from the environment at the logical tick boundaries
(cf. Section~\ref{sec:mapping-logical-time}). In
Figure~\ref{fig:resolution}, assuming reaction R2 performs some large
data computation the length of the tick is determined by R2. Consider
what happens if the external timer generates the \texttt{TIMER} signal,
but the system is still performing computation and the logical tick has
not yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
environment. Thus, even in the absence of the \texttt{suspend}
construct, we are not guaranteed that the program will capture the
external timer signal. In fact, the WCRT needs to be smaller than 1 ms
for the above code snippet to perform as expected. But, this again
implies need for a time analyzable platform even when using external
timers.

In general we observe that the WCRT determines the lowest resolution of
any real-time construct whether it be external timer independent, like
the proposed technique, or external timer dependent, like above. This
analysis is one of the key insights for proposing the solutions
described in this paper.

### 2014-01-31 exps.tex
Updated the paper

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time waits and program sizes are shown in
Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exact(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exact(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exact(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exact(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

\hj{ HRTCS is the human response system described in
	Section~\ref{sec:motivating-example}. HRTCS consists of 2 CDs, but
only one with real-time waits.}

@@

AECS (Access and Environmental Control System) is an application that
controls an intelligent room environment described in~\cite{aecs_ispa}.
AECS has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2
wait statements. The system is designed to measure how long the entrance
door to the room is being opened or to detect unknown entries (i.e.
intruders). For instance, the system triggers an alarm 10 seconds after
detection of an intruder or generates a beep sound when an entrance door
is being opened for the same amount of time. In the original AECS, the
system needed to interacts with two external timers through input/output
signals in order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await} statement.
The program has been modified such that the wait constructs replace
external timers and corresponding interface (I/O) signals, which is
shown in Figure~\ref{delaycode:b}.  With the new wait constructs, these
timers are not necessary and waits can be expressed directly in the
code.

@@

The previous results show that for HRTCS and Printer the compiler was
able to find \textit{d} without relaxing upper bound of wait
statements. Figure~\ref{fig:graph1} shows the accuracy of our results by
comparing differences between programmer specified upper bounds
(Table~\ref{fig:benchlist}) and the actual bounds obtained using our
techniques (Table~\ref{fig:comparison}) for the other examples requiring
relaxation. AECS/CD1 wait specification is not included in
Figure~\ref{fig:graph2}, because although it requires relaxation, the
wait specification (Table~\ref{fig:benchlist}) does not have an upper
bound.

@@

There are two influential factors in determining the relaxation: 1)
execution speed of a target platform and 2) the nature of the program
that runs on this platform. Clock frequency of the processor directly
impacts the value of WCRT. On the other hand, programmer can write a
program such that large data (or control) computations in a single
logical tick increase WCRT of the program.  Assuming we do not change
the architecture of the execution platform (i.e. JOP), the differece
factor can only be reduced by modifying the benchmark programs such that
the critical paths are broken up using \texttt{pause} constructs,
thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
code from AECS whose WCRT is reduced by such technique.

>>>>

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time waits and program sizes are shown in
Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exact(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exact(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exact(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exact(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.112 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

\hj{ HRTCS is the human response system described in
  Section~\ref{sec:motivating-example}. HRTCS consists of 2 CDs, but
  only one with real-time wait.}

@@

AECS (Access and Environmental Control System) is an application that
controls an intelligent room environment described in~\cite{aecs_ispa}.
AECS has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2
wait statements. The system is designed to measure how long the entrance
door to the room is being opened or to detect unknown entries (i.e.
intruders). For instance, the system triggers an alarm 10 seconds after
detection of an intruder or generates a beep sound when an entrance door
is being opened for the same amount of time. In the original AECS, the
system needed to interact with two external timers through input/output
signals in order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await} statement.
The program has been modified such that the wait constructs replace
external timers and corresponding interface (I/O) signals, which is
shown in Figure~\ref{delaycode:b}.  With the new wait constructs, these
timers are not necessary and real-time waits can be expressed directly
in the SystemJ code.

@@

The previous results show that for HRTCS and Printer the compiler was
able to find \textit{d} without relaxing upper bound of wait
statements. Figure~\ref{fig:graph1} shows the accuracy of our results by
comparing between the programmer specified upper bounds
(Table~\ref{fig:benchlist}) and the actual bounds obtained using our
techniques (Table~\ref{fig:comparison}) for the examples requiring
relaxation. AECS/CD1 wait specification is not included in
Figure~\ref{fig:graph2}, because although it requires relaxation, the
wait specification (Table~\ref{fig:benchlist}) does not have an upper
bound.

@@

There are two influential factors in determining the relaxation: 1)
execution speed of a target platform and 2) the nature of the program
that runs on this platform. Clock frequency of the processor directly
impacts the value of WCRT. On the other hand, programmer can write a
program such that large data (or control) computations in a single
logical tick increase WCRT of the program.  Assuming we do not change
the architecture of the execution platform (i.e. JOP), the difference
factor can only be reduced by modifying the benchmark programs such that
the critical paths are broken up using \texttt{pause} constructs,
thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
code from AECS whose WCRT is reduced by such technique.
